---
title: "PRoject"
output:
  word_document: default
  html_document: default
  pdf_document: default
subtitle: 
author: sophie and Diya
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(car)
library(leaps)
myResPlots <- function(model, label){
  
  #Normal quantile plot of studentized residuals
  qqPlot(rstudent(model), pch = 19, main = paste("NQ Plot of Studentized Residuals,", label))
  
  #plot of fitted vs. studentized residuals
  plot(rstudent(model) ~ model$fitted.values, pch = 19, col = 'red', xlab = "Fitted Values", ylab = "Studentized Residuals",
     main = paste("Fits vs. Studentized Residuals,", label))
  abline(h = 0, lwd = 3)
  abline(h = c(2,-2), lty = 2, lwd = 2, col="blue")
  abline(h = c(3,-3), lty = 2, lwd = 2, col="green")

}
```

This assignment uses data from the Anime Recommendation Database 2020 on Kaggle, based on stats on myanimelist.

```{r}
anime <- read.csv("https://github.com/diyamagnetism/Stats230/raw/refs/heads/main/anime.csv")
dim(anime)
names(anime)
head(anime)



```

```{r}

library(tidyr)

anime[anime == "Unknown"] <- NA

genres_split <- strsplit(anime$Genres, ",\\s*")
table(unlist(genres_split))
unique_genres <- sort(unique(unlist(genres_split)))

anime$Genres <- gsub("Hentai|Ecchi|Harem", "Explicit", anime$Genres)
anime$Genres <- gsub("Yuri|Yaoi|Shounen Ai|Shoujo Ai", "LGBT", anime$Genres)
anime$Genres <- gsub("Dementia", "Psychological", anime$Genres)
anime$Genres <- gsub("Demons|Magic|Supernatural|Super Power|Vampire", "Fantasy", anime$Genres)
anime$Genres <- gsub("Space|Mecha", "Sci-Fi", anime$Genres)
anime$Genres <- gsub("Military|Cars|Police|Martial Arts", "Action", anime$Genres)
anime$Genres <- gsub("Samurai", "Historical", anime$Genres)
anime$Genres <- gsub("Parody", "Comedy", anime$Genres)


anime2 <- anime[, c("Name", "Score", "Type", "Episodes", "Aired", "Studios", "Source", "Duration", "Rating", "Popularity", "Members", "Favorites", "Dropped", "Plan.to.Watch", "Score.10", "Score.9", "Score.8", "Score.7", "Score.6", "Score.5", "Score.4", "Score.3", "Score.2", "Score.1", "Completed")]

#anime2 <- separate_longer_delim(anime2, Genres, delim=",")

genres_split <- strsplit(anime$Genres, ",\\s*")
table(unlist(genres_split))
unique_genres <- sort(unique(unlist(genres_split)))


for (genre in unique_genres){
  anime2[[genre]] <- sapply(genres_split, function(x) as.integer(genre %in% x))
}

head(anime2)


```
### Data Cleaning

First, I replaced all values labeled as `"Unknown"` with proper missing values (`NA`) to standardize missing data across the dataset.

## Cleaning Genres 

Originally, each anime’s Genres column contained a comma-separated list of multiple genres. The dataset initially included the following unique genres:

```
Action, Adventure, Cars, Comedy, Dementia, Demons, Drama, Ecchi, Fantasy,
Game, Harem, Hentai, Historical, Horror, Josei, Kids, Magic, Martial Arts,
Mecha, Military, Music, Mystery, Parody, Police, Psychological, Romance, Samurai,
School, Sci-Fi, Seinen, Shoujo, Shoujo Ai, Shounen, Shounen Ai, Slice of Life, Space,
Sports, Super Power, Supernatural, Thriller, Vampire, Yaoi, Yuri
```

To simplify and group related genres, I consolidated them as follows:

* "Hentai", "Ecchi", and "Harem" were grouped into "Explicit".
* "Yuri", "Yaoi", "Shounen Ai", and "Shoujo Ai" were grouped into "LGBT".
* "Dementia" was recategorized as "Psychological".
* "Demons", "Magic", "Supernatural", "Super Power", and "Vampire" were grouped under "Fantasy".
* "Space" and "Mecha" were combined into "Sci-Fi".
* "Military", "Cars", "Police", and "Martial Arts" were grouped as "Action".
* "Samurai" was recategorized as "Historical".
* "Parody" was grouped under "Comedy".

After cleaning and grouping, I transformed the Genres column into multiple boolean (0/1) columns, where each column represents a genre, and each anime is marked `1` if it belongs to that genre and `0` if not. The resulting genre columns are:

```
"Action", "Adventure", "Comedy", "Drama", "Explicit", "Fantasy", "Game",
"Historical", "Horror", "Josei", "Kids", "LGBT", "Music", "Mystery",
"Psychological", "Romance", "School", "Sci-Fi", "Seinen", "Shoujo", "Shounen",
"Slice of Life", "Sports", "Thriller"
```

This format makes genre-based filtering and analysis much easier.


```{r}
anime2 <- anime2[!is.na(anime2$Score), ]

anime2[is.na(anime2$Episodes),]
anime2[is.na(anime2$Episodes)&anime2$Popularity<5000, ][order(anime2$Popularity[is.na(anime2$Episodes)&anime2$Popularity<5000]), ]

anime2[anime2$Name == "One Piece",]$Episodes <- 923

anime2[anime2$Name == "Sazae-san",]$Episodes <- 2771

anime2[anime2$Name == "Boruto: Naruto Next Generations",]$Episodes <- 145

anime2[anime2$Name == "Detective Conan",]$Episodes <- 970

anime2[anime2$Name == "Beastars 2nd Season",]$Episodes <- 12

anime2[anime2$Name == "Crayon Shin-chan",]$Episodes <- 1026

anime2[anime2$Name == "Super Dragon Ball Heroes",]$Episodes <- 20

anime2[anime2$Name == "Holo no Graffiti",]$Episodes <- 42
  
anime2[anime2$Name == "Fate/Grand Order",]$Episodes <- 20

anime2[anime2$Name == "Doraemon (2005)",]$Episodes <- 591

anime2[anime2$Name == "Nintama Rantarou",]$Episodes <- 2169

anime2[anime2$Name == "Ojarumaru",]$Episodes <- 2082

anime2[anime2$Name == "Sore Ike! Anpanman",]$Episodes <- 1587
  
anime2[anime2$Name == "Chibi Maruko-chan (1995)",]$Episodes <- 1468

```
## Missing Episode Counts

I noticed that several popular series had missing values in their Episode Count column. After reviewing the data, I identified three main reasons for this:

1. The series was ongoing at the time of data collection.
2. The series was very long-running, often leading to incomplete entries.
3. The anime lacked reliable data and had multiple missing columns.

To address this:

* I first removed any anime entries with missing Score values, which helped eliminate most of the incomplete and unreliable rows (resolving issue #3).
* Then, for around a dozen well-known long-running series, I manually updated their episode counts using accurate figures from around 2020 (the year this dataset reflects). Examples include One Piece, Detective Conan, Crayon Shin-chan, and others, where I input their correct episode numbers.

This process ensured that the Episode Count column is as complete and accurate as possible, especially for the most significant and popular titles in the dataset.


```{r}
#anime2$Duration clean this

sum(is.na(anime2$Duration))

hours <- as.numeric(gsub("?([0-9]) hr..*", "\\1", anime2$Duration))
hours[is.na(hours)]<-0
minutes<-gsub(".*?([0-9]{2}) min..*", "\\1", anime2$Duration)
minutes<-gsub(".*?([0-9]{1}) min..*", "\\1", minutes)
minutes<-as.numeric(minutes)
minutes[is.na(minutes)]<-0

anime2$EpisodeLength <- 60*hours + minutes
anime2$TotalRuntime <- (anime2$EpisodeLength)*(as.numeric(anime2$Episodes))


```

## Episode Duration Cleaning

The Duration column originally contained strings like `"24 min. per ep."` or `"1 hr. 55 min."`.  
To convert this into numeric values (total minutes per episode), I used `gsub()` to extract the hour and minute components separately.  
Then, I calculated a new column EpisodeLength (in minutes) and another column TotalRuntime by multiplying the episode length by the total number of episodes.

```{r}

anime2$Year = as.integer(gsub(".* ([0-9]{4}).*", "\\1", anime2$Aired))
anime2$Decade = floor((anime2$Year)/10)*10



anime2$Rating <- gsub("?(.*) -.*", "\\1", anime2$Rating)
table(anime2$Rating)



anime2$Source <- gsub(".*manga.*", "Manga", anime2$Source)
anime2$Source <- gsub(".*book.*", "Novel", anime2$Source, ignore.case=TRUE)
anime2$Source <- gsub("Visual novel|Card game", "Game", anime2$Source)
anime2$Source <- gsub("Music|Radio", "Other", anime2$Source)
unique(anime2$Source)

head(anime2)
```
## Year and Decade

From the Aired column (which contained ranges like `"Apr 3, 1998 to Apr 24, 1999"`),  I extracted the first 4-digit year using `gsub()` and stored it in a new column Year. Then, I calculated the Decade by flooring the year to the nearest decade (e.g., 1998 → 1990). The Decade column will be a useful feature for analyzing trends over time —  for example, to compare the popularity, genres, or scores of anime released in different decades.

## Cleaning Rating Categories

The Rating column originally had extra descriptions (e.g., `"PG-13 - Teens 13 or older"`).  I cleaned it using `gsub()` to keep only the core rating (e.g., `"PG-13"`).  This makes the rating categories consistent and easier to print in figures.

## Cleaning Source Material Categories

Similarly to the Genre column, I simplified the Source column:

- Any source mentioning "manga" is categorized as `"Manga"`.
- Sources mentioning "book" are categorized as `"Novel"`.
- `"Visual novel"` and `"Card game"` are grouped as `"Game"`.
- `"Music"` and `"Radio"` are grouped as `"Other"`.


```{r}

anime2$Score <- as.numeric(anime2$Score)
anime2$Episodes <- as.numeric(anime2$Episodes)
anime2$Score.1 <- as.numeric(anime2$Score.1)
anime2$Score.2 <- as.numeric(anime2$Score.2)
anime2$Score.3 <- as.numeric(anime2$Score.3)
anime2$Score.4 <- as.numeric(anime2$Score.4)
anime2$Score.5 <- as.numeric(anime2$Score.5)
anime2$Score.6 <- as.numeric(anime2$Score.6)
anime2$Score.7 <- as.numeric(anime2$Score.7)
anime2$Score.8 <- as.numeric(anime2$Score.8)
anime2$Score.9 <- as.numeric(anime2$Score.9)
anime2$Score.10 <- as.numeric(anime2$Score.10)

anime2
```
```{r}
# labels outliers in a boxplot! use for something
library(TeachingDemos)

library(plyr)
source("https://raw.githubusercontent.com/talgalili/R-code-snippets/master/boxplot.with.outlier.label.r") # Load the function

#par(cex = .7)
#boxplot.with.outlier.label(wb$Guns_per_100, wb$Country, col = "red", main = "Total Guns per 100", cex = .7)

```


```{r}


anime2$logDropped <- log10(anime2$Dropped)
anime2

boxplot(anime2$logDropped ~ anime2$Rating,
        xlab = "Rating",
        ylab = "log Dropped",
        col = c(2:8))


boxplot(anime2$logDropped ~ anime2$Source,
        xlab = "Source",
        ylab = "log Dropped",
        col = c(2:7, 2:7))



#boxplot.with.outlier.label(anime2$logDropped, anime2$Name)

```
```{r}


hist(anime2$Score, 
     main = "Scores of Anime",
     xlab = "Score",
     ylab = "Frequency",
     breaks = 30,
     xlim = c(0,10),
     col = "green")

anime2$logMembers <- log10(anime2$Members)

hist(anime2$logMembers, 
     main = "Members",
     xlab = "Log Number of Members",
     ylab = "Frequency",
     breaks = 50,
     col = "blue")


```
```{r}

par(mfrow = c(4, 6), mar = c(2, 2, 2, 1), oma = c(0, 0, 4, 0))
for (genre in unique_genres) {
  genre_data <- subset(anime2, anime2[[genre]] == 1)
  hist(genre_data$Year, breaks = seq(min(anime2$Year, na.rm = TRUE), max(anime2$Year, na.rm = TRUE), by = 1), main = genre, xlab = "Year", ylab = "Count", border = "skyblue")
}
mtext("Number of Anime by Genre and Year", outer = TRUE, cex = 1.5)

par(mfrow = c(4, 6), mar = c(2, 2, 2, 1), oma = c(0, 0, 4, 0))
for (genre in unique_genres) {
  years <- sort(unique(anime2$Year[!is.na(anime2$Year)]))
  counts_in_genre <- sapply(years, function(y) sum(anime2[[genre]] == 1 & anime2$Year == y, na.rm = TRUE))
  counts_total <- sapply(years, function(y) sum(anime2$Year == y, na.rm = TRUE))
  proportions <- counts_in_genre / counts_total
  
  barplot(proportions, names.arg = years, main = paste(genre), xlab = "Year", ylab = "Proportion", border = "skyblue", las = 2, cex.names = 0.7)
}

mtext("Proportion of Anime by Genre and Year", outer = TRUE, cex = 1.5)

```

```{r}

anime2$logCompleted <- log10(anime2$Completed)



t.test(anime2$Completed - anime2$Members, alternative = "two.sided", conf.level = 0.95, paired = FALSE)   



# t test Completed - Dropped-- on average, is there a difference between members following an anime
# people who have completed a show? apparently so.

```

```{r}
anime2$MGenderMarket <- as.integer(anime2$Shounen + anime2$Seinen)
anime2$MGenderMarket <- gsub(1,1, anime2$MGenderMarket)
anime2$FGenderMarket <- as.integer(anime2$Shoujo + anime2$Josei)
anime2$FGenderMarket <- gsub(1,2, anime2$FGenderMarket)



anime2$FGenderMarket = as.integer(anime2$FGenderMarket)
anime2$MGenderMarket = as.integer(anime2$MGenderMarket)



anime2$GenderMarketing <- anime2$FGenderMarket + anime2$MGenderMarket

anime2$GenderMarketing[anime2$GenderMarketing == 0] <- NA
anime2

```

```{r}


boyscores <- (anime2$Score[anime2$GenderMarketing == 1])
girlscores <- (anime2$Score[anime2$GenderMarketing == 2])

mean(na.omit(boyscores))
mean(na.omit(girlscores))

actualdiff <- mean(na.omit(boyscores)) - mean(na.omit(girlscores))

(actualdiff)



```


```{r}

# permuted sample mean difference between shounen + seinen scores vs shojo + josei scores-- myanimelist is (anecdotally) a male dominated site.
# do shojo and josei routinely get disfavored?

# boyMarketed <- shounen plus seinen
# girlMarketed <- shojo plus josei


set.seed(1)

N <- 10000

diffvals <- rep(NA, N)

for (i in 1:N) {
  fakeGM <- sample(anime2$GenderMarketing)  # default is replace = FALSE
  diffvals[i] <- mean(na.omit(anime2$Score[fakeGM == 1])) -  mean(na.omit(anime2$Score[fakeGM == 2]))
}





```

```{r}

#Make histogram of permuted mean differences
hist(diffvals, col = "magenta", main = "Permuted Sample Means Difference in Scores\n Between Anime Marketed Towards Men vs. Women", xlab = "Difference in Scores (M - F)", breaks = 50, 
     xlim = c(-0.25,0.25))
abline(v = actualdiff, col = "blue", lwd = 3)
text(actualdiff - 0.03, 300 , paste("Actual Diff in Means =", round(actualdiff,2)),srt = 90)
```

```{r}


years <- sort(unique(anime2$Year[!is.na(anime2$Year)]))
counts_lgbt <- sapply(years, function(y) sum(anime2$LGBT == 1 & anime2$Year == y, na.rm = TRUE))
counts_total <- sapply(years, function(y) sum(anime2$Year == y, na.rm = TRUE))
pct_lgbt <- counts_lgbt / counts_total * 100
  
barplot(pct_lgbt, names.arg = years, xlab = "Year", ylab = "Proportion", col = "skyblue", las = 2, cex.names = 0.7)


years_before <- years[years >= 1981 & years <= 2000]
years_after <- years[years > 2000]

pct_before <- pct_lgbt[years %in% years_before]
pct_after <- pct_lgbt[years %in% years_after]



N <- 10000
diff <- rep(NA, N)


for (i in 1:N) {
  sample_before <- sample(pct_before, replace = TRUE)
  sample_after <- sample(pct_after, replace = TRUE)
  diff[i] <- mean(sample_after) - mean(sample_before)
}

mean_diff <- mean(diff)
ci <- quantile(diff, c(0.025, 0.975))

paste(ci)

hist(diff, breaks = 50, main = "Bootstrap Distribution of Difference in Yearly LGBT Percentage", 
     xlab = "Difference in Proportions (After - Before)", col = "skyblue")
abline(v = ci, col = "red", lty = 2)
abline(v = mean_diff, col = "blue", lwd = 2)





```


```{r}
# Multiple Regression with Score 1-10



anime2

anime3 <- anime2[, c("Score", "TotalRuntime", "logMembers", "Score.1", "Score.5", "Score.10", "logDropped", "logCompleted")]

anime3

anime



```


```{r}


anime3$logDropped[!is.finite(anime3$logDropped)] <- NA 
anime3$logCompleted[!is.finite(anime3$logCompleted)] <- NA 

anime3$logScore.1 <- log(anime3$Score.1)
anime3$logScore.5 <- log(anime3$Score.5)
anime3$logScore.10 <- log(anime3$Score.10)

#Make object called WB4 which has complete data for variables we care about
anime4 <- na.omit(anime3[, c("Score", "TotalRuntime", "logMembers", "logScore.1", "logScore.5", "logScore.10", "logDropped", "logCompleted")])





anime4 <- anime4[complete.cases(anime4), ]

anime4





```




```{r}

library(corrplot)

```


```{r}
# the questionmarks occur when there is not enough variance between variables



sigcorr <- cor.mtest(anime4, conf.level = .95)
corrplot.mixed(cor(anime4), lower.col="black", upper = "ellipse", tl.col = "black", number.cex=.7, 
                tl.pos = "lt", tl.cex=.7, p.mat = sigcorr$p, sig.level = .05)
```
```{r}
library(PerformanceAnalytics)
```

```{r}
# lots of nonlinearity
chart.Correlation(anime4, histogram = TRUE, pch = 19)




```

```{r}
c("Score", "TotalRuntime", "logMembers", "logScore.1", "logScore.5", "logScore.10", "logDropped", "logCompleted")

```

We cut logCompleted for colinearity reasons, as well as score and total runtime for lack of linearity


```{r}
#model with all predictors
lm1 <- lm(logMembers ~ logScore.1 + logScore.5 + logScore.10 + logDropped , data = anime4)
summary(lm1)

  
myResPlots(lm1, "Model for Score")

```

We remove logScore.1 for having a high p value in tryng to predict logMembers

```{r}
#model with all predictors
lm2 <- lm(logMembers ~ logScore.5 + logScore.10 + logDropped , data = anime4)
summary(lm2)

  
myResPlots(lm2, "Model for Members")

```




```{r}
trans1 <- boxCox(lm2)
```
```{r}

variable_list <-  c("logDropped", "Rating", "Score.1", "Dropped", "logMembers", "Score")

anime5 <- anime2[, variable_list]


anime5$logScore.1 <- log(anime5$Score)

anime5$logDropped[!is.finite(anime5$logDropped)] <- NA 

anime4 <- na.omit(anime5[, variable_list])



plot(logDropped ~ Score, data = anime5)


m5 <- lm(logDropped ~ Score*Rating, data = anime5)
#Again, get overall test of significance of terms
myResPlots(m5, "Dropped")

Anova(m5, type = 3)

summary(m5)
```
```{r}

plot(logDropped ~ Score, data = anime5, col = factor(Rating), pch = 16, cex = .5, xlim = c(0,10), ylim = c(0,10))
legend("topleft", col = 1:6, legend = levels(factor(anime5$Rating)), pch = 16)
coefs <- coef(m5)
#Again, pay CLOSE attention to which coefficient is which!!!!
round(coefs,4)


abline(a = coefs[1], b = coefs[2], col = "black", lwd = 3)
for (i in 3:7){
  abline(a = coefs[1] + coefs[i], b = coefs[2] + coefs[i+4], col = (i-1), lwd = 3)
}
```

```{r}

```


