---
title: "PRoject"
output:
  word_document: default
  html_document: default
  pdf_document: default
subtitle: 
author: sophie and Diya
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(car)
library(leaps)
myResPlots <- function(model, label){
  
  #Normal quantile plot of studentized residuals
  qqPlot(rstudent(model), pch = 19, main = paste("NQ Plot of Studentized Residuals,", label))
  
  #plot of fitted vs. studentized residuals
  plot(rstudent(model) ~ model$fitted.values, pch = 19, col = 'red', xlab = "Fitted Values", ylab = "Studentized Residuals",
     main = paste("Fits vs. Studentized Residuals,", label))
  abline(h = 0, lwd = 3)
  abline(h = c(2,-2), lty = 2, lwd = 2, col="blue")
  abline(h = c(3,-3), lty = 2, lwd = 2, col="green")

}
```

This assignment uses data from the Anime Recommendation Database 2020 on Kaggle, based on stats on myanimelist.

```{r}
anime <- read.csv("https://github.com/diyamagnetism/Stats230/raw/refs/heads/main/anime.csv")
dim(anime)
names(anime)
head(anime)



```

```{r}

library(tidyr)

anime[anime == "Unknown"] <- NA

genres_split <- strsplit(anime$Genres, ",\\s*")
table(unlist(genres_split))
unique_genres <- sort(unique(unlist(genres_split)))

anime$Genres <- gsub("Hentai|Ecchi|Harem", "Explicit", anime$Genres)
anime$Genres <- gsub("Yuri|Yaoi|Shounen Ai|Shoujo Ai", "LGBT", anime$Genres)
anime$Genres <- gsub("Dementia", "Psychological", anime$Genres)
anime$Genres <- gsub("Demons|Magic|Supernatural|Super Power|Vampire", "Fantasy", anime$Genres)
anime$Genres <- gsub("Space|Mecha", "Sci-Fi", anime$Genres)
anime$Genres <- gsub("Military|Cars|Police|Martial Arts", "Action", anime$Genres)
anime$Genres <- gsub("Samurai", "Historical", anime$Genres)
anime$Genres <- gsub("Parody", "Comedy", anime$Genres)


anime2 <- anime[, c("Name", "Score", "Type", "Episodes", "Aired", "Studios", "Source", "Duration", "Rating", "Popularity", "Members", "Favorites", "Dropped", "Plan.to.Watch", "Score.10", "Score.9", "Score.8", "Score.7", "Score.6", "Score.5", "Score.4", "Score.3", "Score.2", "Score.1", "Completed")]

#anime2 <- separate_longer_delim(anime2, Genres, delim=",")

genres_split <- strsplit(anime$Genres, ",\\s*")
table(unlist(genres_split))
unique_genres <- sort(unique(unlist(genres_split)))


for (genre in unique_genres){
  anime2[[genre]] <- sapply(genres_split, function(x) as.integer(genre %in% x))
}

head(anime2)


```
### Data Cleaning

First, I replaced all values labeled as `"Unknown"` with proper missing values (`NA`) to standardize missing data across the dataset.

## Cleaning Genres 

Originally, each anime’s Genres column contained a comma-separated list of multiple genres. The dataset initially included the following unique genres:

```
Action, Adventure, Cars, Comedy, Dementia, Demons, Drama, Ecchi, Fantasy,
Game, Harem, Hentai, Historical, Horror, Josei, Kids, Magic, Martial Arts,
Mecha, Military, Music, Mystery, Parody, Police, Psychological, Romance, Samurai,
School, Sci-Fi, Seinen, Shoujo, Shoujo Ai, Shounen, Shounen Ai, Slice of Life, Space,
Sports, Super Power, Supernatural, Thriller, Vampire, Yaoi, Yuri
```

To simplify and group related genres, I consolidated them as follows:

* "Hentai", "Ecchi", and "Harem" were grouped into "Explicit".
* "Yuri", "Yaoi", "Shounen Ai", and "Shoujo Ai" were grouped into "LGBT".
* "Dementia" was recategorized as "Psychological".
* "Demons", "Magic", "Supernatural", "Super Power", and "Vampire" were grouped under "Fantasy".
* "Space" and "Mecha" were combined into "Sci-Fi".
* "Military", "Cars", "Police", and "Martial Arts" were grouped as "Action".
* "Samurai" was recategorized as "Historical".
* "Parody" was grouped under "Comedy".

After cleaning and grouping, I transformed the Genres column into multiple boolean (0/1) columns, where each column represents a genre, and each anime is marked `1` if it belongs to that genre and `0` if not. The resulting genre columns are:

```
"Action", "Adventure", "Comedy", "Drama", "Explicit", "Fantasy", "Game",
"Historical", "Horror", "Josei", "Kids", "LGBT", "Music", "Mystery",
"Psychological", "Romance", "School", "Sci-Fi", "Seinen", "Shoujo", "Shounen",
"Slice of Life", "Sports", "Thriller"
```

This format makes genre-based filtering and analysis much easier.


```{r}
anime2 <- anime2[!is.na(anime2$Score), ]

anime2[is.na(anime2$Episodes),]
anime2[is.na(anime2$Episodes)&anime2$Popularity<5000, ][order(anime2$Popularity[is.na(anime2$Episodes)&anime2$Popularity<5000]), ]

anime2[anime2$Name == "One Piece",]$Episodes <- 923

anime2[anime2$Name == "Sazae-san",]$Episodes <- 2771

anime2[anime2$Name == "Boruto: Naruto Next Generations",]$Episodes <- 145

anime2[anime2$Name == "Detective Conan",]$Episodes <- 970

anime2[anime2$Name == "Beastars 2nd Season",]$Episodes <- 12

anime2[anime2$Name == "Crayon Shin-chan",]$Episodes <- 1026

anime2[anime2$Name == "Super Dragon Ball Heroes",]$Episodes <- 20

anime2[anime2$Name == "Holo no Graffiti",]$Episodes <- 42
  
anime2[anime2$Name == "Fate/Grand Order",]$Episodes <- 20

anime2[anime2$Name == "Doraemon (2005)",]$Episodes <- 591

anime2[anime2$Name == "Nintama Rantarou",]$Episodes <- 2169

anime2[anime2$Name == "Ojarumaru",]$Episodes <- 2082

anime2[anime2$Name == "Sore Ike! Anpanman",]$Episodes <- 1587
  
anime2[anime2$Name == "Chibi Maruko-chan (1995)",]$Episodes <- 1468

```
## Missing Episode Counts

I noticed that several popular series had missing values in their Episode Count column. After reviewing the data, I identified three main reasons for this:

1. The series was ongoing at the time of data collection.
2. The series was very long-running, often leading to incomplete entries.
3. The anime lacked reliable data and had multiple missing columns.

To address this:

* I first removed any anime entries with missing Score values, which helped eliminate most of the incomplete and unreliable rows (resolving issue #3).
* Then, for around a dozen well-known long-running series, I manually updated their episode counts using accurate figures from around 2020 (the year this dataset reflects). Examples include One Piece, Detective Conan, Crayon Shin-chan, and others, where I input their correct episode numbers.

This process ensured that the Episode Count column is as complete and accurate as possible, especially for the most significant and popular titles in the dataset.


```{r}
#anime2$Duration clean this

sum(is.na(anime2$Duration))

hours <- as.numeric(gsub("?([0-9]) hr..*", "\\1", anime2$Duration))
hours[is.na(hours)]<-0
minutes<-gsub(".*?([0-9]{2}) min..*", "\\1", anime2$Duration)
minutes<-gsub(".*?([0-9]{1}) min..*", "\\1", minutes)
minutes<-as.numeric(minutes)
minutes[is.na(minutes)]<-0

anime2$EpisodeLength <- 60*hours + minutes

anime2$Episodes <- as.numeric(anime2$Episodes)

anime2$TotalRuntime <- anime2$Duration * anime2$Episodes


```

## Episode Duration Cleaning

The Duration column originally contained strings like `"24 min. per ep."` or `"1 hr. 55 min."`.  
To convert this into numeric values (total minutes per episode), I used `gsub()` to extract the hour and minute components separately.  
Then, I calculated a new column EpisodeLength (in minutes) and another column TotalRuntime by multiplying the episode length by the total number of episodes.

```{r}

anime2$Year = as.integer(gsub(".* ([0-9]{4}).*", "\\1", anime2$Aired))
anime2$Decade = floor((anime2$Year)/10)*10



anime2$Rating <- gsub("?(.*) -.*", "\\1", anime2$Rating)
table(anime2$Rating)



anime2$Source <- gsub(".*manga.*", "Manga", anime2$Source)
anime2$Source <- gsub(".*book.*", "Novel", anime2$Source, ignore.case=TRUE)
anime2$Source <- gsub("Visual novel|Card game", "Game", anime2$Source)
anime2$Source <- gsub("Music|Radio", "Other", anime2$Source)
unique(anime2$Source)

head(anime2)
```
## Year and Decade

From the Aired column (which contained ranges like `"Apr 3, 1998 to Apr 24, 1999"`),  I extracted the first 4-digit year using `gsub()` and stored it in a new column Year. Then, I calculated the Decade by flooring the year to the nearest decade (e.g., 1998 → 1990). The Decade column will be a useful feature for analyzing trends over time —  for example, to compare the popularity, genres, or scores of anime released in different decades.

## Cleaning Rating Categories

The Rating column originally had extra descriptions (e.g., `"PG-13 - Teens 13 or older"`).  I cleaned it using `gsub()` to keep only the core rating (e.g., `"PG-13"`).  This makes the rating categories consistent and easier to print in figures.

## Cleaning Source Material Categories

Similarly to the Genre column, I simplified the Source column:

- Any source mentioning "manga" is categorized as `"Manga"`.
- Sources mentioning "book" are categorized as `"Novel"`.
- `"Visual novel"` and `"Card game"` are grouped as `"Game"`.
- `"Music"` and `"Radio"` are grouped as `"Other"`.

## Making Scores/Episodes Numeric

Converted all the Score.#, Score, and Episodes columns into numerics


```{r}

anime2$Score <- as.numeric(anime2$Score)
anime2$Episodes <- as.numeric(anime2$Episodes)
anime2$Score.1 <- as.numeric(anime2$Score.1)
anime2$Score.2 <- as.numeric(anime2$Score.2)
anime2$Score.3 <- as.numeric(anime2$Score.3)
anime2$Score.4 <- as.numeric(anime2$Score.4)
anime2$Score.5 <- as.numeric(anime2$Score.5)
anime2$Score.6 <- as.numeric(anime2$Score.6)
anime2$Score.7 <- as.numeric(anime2$Score.7)
anime2$Score.8 <- as.numeric(anime2$Score.8)
anime2$Score.9 <- as.numeric(anime2$Score.9)
anime2$Score.10 <- as.numeric(anime2$Score.10)


anime2$logCompleted <- log10(anime2$Completed)
anime2$logDropped <- log10(anime2$Dropped)
anime2$logMembers <- log10(anime2$Members)
anime2$logDropped[!is.finite(anime2$logDropped)] <- NA 
anime2$logCompleted[!is.finite(anime2$logCompleted)] <- NA 
anime2$logMembers[!is.finite(anime2$logMembers)] <- NA 

anime2
```

# Data Visualization

## Boxplots of Dropped Animes

We then examine how ratings (G to Rx) and source material affect how many people drop a particular anime.



```{r}
# labels outliers in a boxplot! use for something
library(TeachingDemos)

library(plyr)
source("https://raw.githubusercontent.com/talgalili/R-code-snippets/master/boxplot.with.outlier.label.r") # Load the function

```


```{r}


anime2$logDropped <- log10(anime2$Dropped)
anime2$logDropped[!is.finite(anime2$logDropped)] <- NA 


anime2

boxplot(anime2$logDropped ~ anime2$Rating,
        xlab = "Rating",
        ylab = "log Dropped",
        col = c(2:8))


boxplot(anime2$logDropped ~ anime2$Source,
        xlab = "Source",
        ylab = "log Dropped",
        col = c(2:7, 2:7))




```
Among ratings, R rated animes had the highest median number of drops. G, PG, and Rx rated anime are heavily right skewed, even after a log transformation. R rated animes also had the highest interquartile range. Rx had the lowest range.

As for source material, light novels have the highest median number of drops. However, original and game based animes have the highest number of high drop outliars

## Histograms

The scores of anime appear to be normally distributed, while the log number of Members, or followers of an anime are right skewed. This makes sense because a select few shows are wildly popular.

```{r}


hist(anime2$Score, 
     main = "Scores of Anime",
     xlab = "Score",
     ylab = "Frequency",
     breaks = 30,
     xlim = c(0,10),
     col = "green")



hist(anime2$logMembers, 
     main = "Members",
     xlab = "Log Number of Members",
     ylab = "Frequency",
     breaks = 50,
     col = "blue",
     xlim = c(0,7))


```

# T-test

We then performed a two sided t-test. We tested whether the number of people who have completed an anime
are higher than the number of people who drop an anime. The null hypothesis was that there was no difference between the mean number of people who completed vs dropped an anime. The alternative hypothesis was that there was a difference. We obtained a p value of 2.2e-16 under a 95% confidence interval. The mean difference was 29563 people


```{r}


t.test(anime2$Completed - anime2$Dropped, alternative = "two.sided", conf.level = 0.95, paired = FALSE)   
t.test(anime2$logCompleted - anime2$logDropped, alternative = "two.sided", conf.level = 0.95)  


# t test Completed - Dropped-- on average, is there a difference between members following an anime
# people who have completed a show? apparently so.

```
## Permutation test

MyAnimeList and many other online anime communities are very male-dominated. I was curious to see if the scores of anime marketed towards men were significantly higher than anime marketed towards women. 

I examined this through a permutation test. I created a column titled 'GenderMarketing' and assigned a numeric value to men vs women marketed animes. Male marketed animes were obtained from the 'Shounen' and 'Seinen' columns (teenage vs adult men), wheras women marketed animes were obtained from 'Shoujo' and 'Josei' (teenage vs adult women).

We used a seed value of 1 and had an N of 10,000.


```{r}


anime2$MGenderMarket <- as.integer(anime2$Shounen + anime2$Seinen)
anime2$MGenderMarket <- gsub(1,1, anime2$MGenderMarket)
anime2$FGenderMarket <- as.integer(anime2$Shoujo + anime2$Josei)
anime2$FGenderMarket <- gsub(1,2, anime2$FGenderMarket)



anime2$FGenderMarket = as.integer(anime2$FGenderMarket)
anime2$MGenderMarket = as.integer(anime2$MGenderMarket)



anime2$GenderMarketing <- anime2$FGenderMarket + anime2$MGenderMarket

anime2$GenderMarketing[anime2$GenderMarketing == 0] <- NA
anime2


```

```{r}


boyscores <- (anime2$Score[anime2$GenderMarketing == 1])
girlscores <- (anime2$Score[anime2$GenderMarketing == 2])

mean(na.omit(boyscores))
mean(na.omit(girlscores))

actualdiff <- mean(na.omit(boyscores)) - mean(na.omit(girlscores))

(actualdiff)



```


```{r}

# permuted sample mean difference between shounen + seinen scores vs shojo + josei scores-- myanimelist is (anecdotally) a male dominated site.
# do shojo and josei routinely get disfavored?

# boyMarketed <- shounen plus seinen
# girlMarketed <- shojo plus josei


set.seed(1)

N <- 10000

diffvals <- rep(NA, N)

for (i in 1:N) {
  fakeGM <- sample(anime2$GenderMarketing)  # default is replace = FALSE
  diffvals[i] <- mean(na.omit(anime2$Score[fakeGM == 1])) -  mean(na.omit(anime2$Score[fakeGM == 2]))
}





```



```{r}

#Make histogram of permuted mean differences
hist(diffvals, col = "magenta", main = "Permuted Sample Means Difference in Scores\n Between Anime Marketed Towards Men vs. Women", xlab = "Difference in Scores (M - F)", breaks = 50, 
     xlim = c(-0.25,0.25))
abline(v = actualdiff, col = "blue", lwd = 3)
text(actualdiff - 0.03, 300 , paste("Actual Diff in Means =", round(actualdiff,2)),srt = 90)
```

If there was no difference between scoring of anime marketed towards men vs. women, we would expect the actual difference in means to lie within the normal distribution centered at zero. Instead, it is statistically significant, with the difference in mean scores far from the null hypothesis distribution. The actual difference in mean scores was 0.18.


# Multiple Regression

We then performed a multiple regression to find out what predicts the logarithmic number of members following a particular anime. We looked at the following columns:

- 'Score' : Average Score of anime out of 10
- 'TotalRuntime' : Runtime as calculated by duration of episode times number of episodes
- 'Score.1' : number of people who gave the anime a score of 1
- 'Score.5': number of people who gave the anime a score of 5
- 'Score.10' : number of people who gave the anime a score of 10
- 'logDropped' : log 10 of people who dropped an anime
- 'logCompleted' : log 10 of people who completed an anime

These are all continuous variables.


```{r}
# Multiple Regression with Score 1-10


anime3 <- anime2[, c("Score", "TotalRuntime", "logMembers", "Score.1", "Score.5", "Score.10", "logDropped", "logCompleted")]


anime3


```




```{r}


anime3$logDropped[!is.finite(anime3$logDropped)] <- NA 
anime3$logCompleted[!is.finite(anime3$logCompleted)] <- NA 

anime3$logScore.1 <- log(anime3$Score.1)
anime3$logScore.5 <- log(anime3$Score.5)
anime3$logScore.10 <- log(anime3$Score.10)

#Make object called WB4 which has complete data for variables we care about
anime4 <- na.omit(anime3[, c("Score", "TotalRuntime", "logMembers", "logScore.1", "logScore.5", "logScore.10", "logDropped", "logCompleted")])





anime4 <- anime4[complete.cases(anime4), ]

anime4





```

We produced a correlation plot and correlation chart of the variables of interest. According to the correlation plot, there is a strong positive correlation between 

- logMembers and logScore.5
- logMembers and logScore.10
- logMembers and logDropped 
- logMembers and logCompleted
- logMembers and Score



```{r}

library(corrplot)

```


```{r}
# the questionmarks occur when there is not enough variance between variables



sigcorr <- cor.mtest(anime4, conf.level = .95)
corrplot.mixed(cor(anime4), lower.col="black", upper = "ellipse", tl.col = "black", number.cex=.7, 
                tl.pos = "lt", tl.cex=.7, p.mat = sigcorr$p, sig.level = .05)
```
```{r}
library(PerformanceAnalytics)
```

Meanwhile, the correlation chart indicates some of these high correlations are misleading and due to non-linearity.
Before proceeding to multiple linear regression, we cut:

- logCompleted for colinearity reasons
- score for lack of linearity
- total runtime for lack of linearity


```{r}
# lots of nonlinearity
chart.Correlation(anime4, histogram = TRUE, pch = 19)

```

```{r}
c("Score", "TotalRuntime", "logMembers", "logScore.1", "logScore.5", "logScore.10", "logDropped", "logCompleted")

```




```{r}
#model with all predictors
lm1 <- lm(logMembers ~ logScore.1 + logScore.5 + logScore.10 + logDropped , data = anime4)
summary(lm1)

  
myResPlots(lm1, "Model for Score")

```

We remove logScore.1 for having a high p value in trying to predict logMembers.
There wasn't that much left to do in this regression analysis. 


```{r}
#model with all predictors
lm2 <- lm(logMembers ~ logScore.5 + logScore.10 + logDropped , data = anime4)
summary(lm2)

  
myResPlots(lm2, "Model for Members")

```

We see that the distribution is relatively normal except for extremely high values. Otherwise the residuals fall within the normal bounds. However, there is some heteroskacdicity in our plot

Our model is close to 1 on box cox, indicating it is relatively normal.

```{r}
trans1 <- boxCox(lm2)
```




```{r}

variable_list <-  c("logDropped", "Rating", "Score.1", "Dropped", "logMembers", "Score")

anime5 <- anime2[, variable_list]


anime5$logScore.1 <- log(anime5$Score)

anime5$logDropped[!is.finite(anime5$logDropped)] <- NA 

anime4 <- na.omit(anime5[, variable_list])



plot(logDropped ~ Score, data = anime5)


m5 <- lm(logDropped ~ Score*Rating, data = anime5)
#Again, get overall test of significance of terms
myResPlots(m5, "Dropped")

Anova(m5, type = 3)

summary(m5)
```
```{r}

plot(logDropped ~ Score, data = anime5, col = factor(Rating), pch = 16, cex = .5, xlim = c(0,10), ylim = c(0,10))
legend("topleft", col = 1:6, legend = levels(factor(anime5$Rating)), pch = 16)
coefs <- coef(m5)
#Again, pay CLOSE attention to which coefficient is which!!!!
round(coefs,4)


abline(a = coefs[1], b = coefs[2], col = "black", lwd = 3)
for (i in 3:7){
  abline(a = coefs[1] + coefs[i], b = coefs[2] + coefs[i+4], col = (i-1), lwd = 3)
}
```

```{r}

```


